### 2021-03-12 paper.tex

%
% Angol nyelvű szakdolgozat minta az Eszterházy Károly Egyetem hallgatóinak.
%

\documentclass[
% without options: one-sided printing, online version
% twoside,       % two-sided printing
]{thesis-ekf}
\setcounter{secnumdepth}{5}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{natbib}
\usepackage{lipsum}
\usepackage{pgfplots}
\usetikzlibrary{intersections}
\usepackage{csquotes}
\renewcommand{\mkbegdispquote}[2]{\itshape}
\usepackage{listings}
\usepackage{color}
\usepackage{subcaption}
\usepackage[colorinlistoftodos]{todonotes}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,
	language=Python,
	aboveskip=3mm,
	belowskip=3mm,
	showstringspaces=false,
	columns=flexible,
	basicstyle={\small\ttfamily},
	numbers=none,
	numberstyle=\tiny\color{gray},
	keywordstyle=\color{blue},
	commentstyle=\color{dkgreen},
	stringstyle=\color{mauve},
	breaklines=true,
	breakatwhitespace=true,
	tabsize=3
}
\usepackage{graphicx}
\graphicspath{{fig/}}

\begin{document}

\institute{Eszterhazy Karoly University}
\title{Machine Learning Model for Agricultural Data Applications}
\author{Adeyemo Adedeji Charles\\Computer Science BSc}

\supervisor{Bence Bogdandy\\Associate Professor
		\\bogdandy.bence@uni-eszterhazy.hu}
\city{Eger}
\date{2020/21}
\maketitle

\begin{abstract}
	Prediction of economical growth is a complex task which is essential for planning sustainable economy.
	Economy has a wide range of indicators that are monitored and recorded by governments and international organizations, such as, the Food and Agriculture Organization.
	Agricultural data set of the Food and Agriculture Organization has records on production of crops and other agricultural products whose production strongly relates to the Gross Domestic Product of many countries.
	Assuming that total crop production and agricultural economy growth are highly related, the production of crops and total value of income from agriculture can be learnt by sequential machine learning models.
	Recurrent Neural Networks excel in learning time series and sequential data. 
	This thesis presents experimental results on training various recurrent neural networks for modeling the changes of Agricultural and Gross Domestic Products.  
	Our experimental results showed that the models could achieve \textbf{85}\% accuracy. 	
\end{abstract}

\chapter*{Dedication}

To my mentor and supervisor, Bogdandy Bence without whom this dream would have remained a dream.

\tableofcontents

\chapter*{Introduction}

\todo{This is has to be concise. What is the purpose of my Thesis? What is the aim and the Objectives
for researching this particular area}

Intelligence is the ability to process information to inform future decisions. The field of Artificial Intelligence is the field which focuses on building algorithm models (in this case, Artificial Intelligence \cite{black2009books} Algorithms) that can do this as well - that is - process information to inform future decisions.

Machine Learning itself is a subset of AI that focuses on teaching a machine how to do this without being explicitly programmed.

Deep Learning is a subset of Machine Learning which takes this idea even further. It allows us to automatically extract the useful pieces of information needed to inform these future predictions.

Availability of localized and external data can help to practice *Precision Agriculture*

 Like all other neural networks, deep-learning models don't take as input raw text:
 they only work with numeric tensors. Vectorizing text is the process of transforming text
 into numeric tensors. This can be done in multiple ways

\chapter{Related Works}
\todo{Also known as Literature. Concise. Here, I have to create a diagram that shows my thought process.}

\input{chapters/relatedWorks}

\chapter{Methods}
\input{chapters/methods}

\chapter{Implementation}
\input{chapters/implementation}

\chapter{Discussion}
\input{chapters/discussion}

\chapter{Summary}
\input{chapters/summary}

\bibliographystyle{plain}
\bibliography{reference}
\end{document}